### 2019-07-14 pagtit.tex
changed the title

\begin{titlepage}
  \begin{center}
  \centering{
    \textsc{\Large \sffamily {Universidade Federal do Rio Grande do Sul}}\\
    \textsc{\Large \sffamily {Programa de Pós-Graduação em Física}}\\
    \vspace{0.15cm}
    \textsc{\large  Doctoral Thesis}}\\
\end{center}

\vspace{6cm}

\vfill
\begin{center}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\setcounter{footnote}{2}
{\Large \sffamily \textbf{Time evolution of weakly turbulent processes in
    the presence of collisional interactions in solar plasmas}\footnote[1]{
    Project financed by the by the Conselho Nacional de
    Desenvolvimento Científico e Tecnológico (CNPq) and the
    Coordenação de Aperfeiçoamento Pessoal de Nível Superior (CAPES).}}\\
{\large \sffamily (Evolução temporal de processos fracamente turbulentos
  na presença de efeitos colisionais em plasmas astrofísicos)}\\
\vspace{1.5cm}
\Large Sabrina Tigik Ferrão\\
\end{center}
\vspace{3cm}
\vfill
\hfill
\parbox{.5\textwidth}{Doctoral Thesis prepared under the supervision of
  Professor Luiz Fernando Ziebell, presented to the Physics Graduation
  Program of the Instituto de F\'{\i}sica at UFRGS, in partial fullfillment
  of the requirements for obtaining the title of Doctor in Sciences.}\\
\vfill
\vspace{1cm}
\begin{center}
  { Porto Alegre\\ \today}
\vspace{0.5cm}
\renewcommand{\thefootnote}{\arabic{footnote}}
\end{center}

\end{titlepage}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: