### 2019-11-12 ty.tex
*** empty log message ***

%\newpage
\pagestyle{empty}
\chapter*{Acknowledgements}
% \begin{center}
% {\centering{\chapter*{Acknowledgements}}}
% \end{center}
%\begin{quote}
To my supervisor Luiz Fernando Ziebell, whose guidance, encouragement,
and patience (tons of) made this work possible. Particularly, I would
like to state my sincere appreciation for the trusting he always
demonstrated to have in my capabilities for developing high-quality
research and for his availability to help me when I needed. Professor,
I am grateful for all these years of learning and fruitful work that
are culminating in this doctoral thesis.

To Peter Yoon, for the all the collaboration and for receiving me in
the University of Maryland. Peter, I also want to thank you for all
the encouragement in this beginning of career. That meant a lot to me.

To Dr. Adolfo Viñas who dedicated a part of his time to teach me PIC
simulation and discuss the subtleties of kinetic theory of plasmas.
Dr. Viñas, thank you very much for your time and your availability.

To my friends that never let me down. Nicole for the companionship, life

To my family for the patience, support and for always stand by me even
when I just needed to be alone to do my work. Gabriel,

%\end{quote}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: