### 2019-11-25 ty.tex
Acknowledgments

\newpage
\pagestyle{empty}
% {\centering{\textbf{\large Acknowledgments}}
% \chapter*{Acknowledgements}
\begin{center}
  \textbf{\large Acknowledgments}
% {\centering{\chapter*{Acknowledgments}}}
\end{center}
% \begin{quote}

First of all, I acknowledge the dedication of my supervisor, Luiz Fernando
Ziebell, whose guidance, encouragement, and patience (tons of) made this
work possible. Further, I would like to state my sincere appreciation for
the trusting he always demonstrated to have in my capabilities for developing
high-quality research and for his availability to discuss my work, giving me
insightful advice. % \emph{Professor, I am grateful for all these years of
  % learning and fruitful work that culminated in this doctoral thesis.}

In the academic scope, I would like to thank Peter Yoon for the years of
productive collaboration and for receiving me with great hospitality during
my stay at the University of Maryland.
% \emph{Peter, thank you very much for
% all the encouragement you gave me. It meant a lot to me.}
I also want to
express my earnest appreciation to Dr.~Adolfo Viñas, who dedicated a part of
his time to teach me PIC simulation and discuss the subtleties of the kinetic
theory of plasmas during my time in the University of Maryland.
% \emph{Dr.~Viñas, thank you very much for your time, availability, and guidance.}
As special
mentions, I would like to thank Jeferson Arenzon, Renato Pakter, and Rudi
Gaelzer for their willingness to help me with my postdoc applications when I
randomly appeared at their offices.

From the Graduate Program in Physics, I am grateful to Liane Thier Ruschel for
the kind and efficient assistance in dealing with bureaucratic procedures and
the most diverse document requests. I also thank Mari Nunes for her patience
and goodwill in dealing with everyday trouble of grad students.

On the personal side, I thank Felipe for his comprehension, patience, support
and encouragement during my undergrad and graduate years. I am also grateful
to Gabriel for his love and refined sense of humor that cheers me up even in
the saddest days.

Agradeço especialmente à minha mãe, Diná, por ter me criado para ser uma
mulher livre e a não aceitar qualquer papel na sociedade que não seja o
que eu considere o melhor para mim. \emph{Mãe, essa tese é dedicada a ti.
  Sem teu exemplo, eu não estaria onde estou. Obrigada por tudo!}

From my friends and colleagues from UFRGS, I am grateful to Nicole for the
companionship that ranges from hard-working weekends to life celebrations.
\emph{Nicole, grad school would be terribly boring without you.} I also thank
Amanda for showing me how to be prepared for life, Demétrius for the cheerful
annoyance, and Vinícius for the light talks and for sharing his delicious
bread. A special thanks to Thales, my example of plasma physicist, and to
Fernanda, whose intelligence and dedication will always be my inspiration. 
Special mentions to colleagues that were part of this journey and deserved
to be remembered: Alexandre, Gustavo, João e Larissa.

From Maryland, I am especially grateful to Jacek for the infinite talks, solid
companionship and (mutual) support.\emph{ Jacek, you made my time in CP much
  better. Dziękuję bardzo!}

%\end{quote}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: